---
title: "Project 2"
author: "Group 7 - Dustin Dembeck, Elizabeth Walter, Brendan Poe, Cong Fu"
date: "11/7/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1
### Data Cleaning & Summarizing

```{r 1}
library(tidyverse)    # For ggplot, dplyr, and friends
library(countrycode)  # For dealing with country names, abbreviations, and codes
library(lubridate)    # For dealing with dates
library(scico)      # for color palettes
```

```{r 2}
refugees_raw <- read_csv("C:/Users/walte/Desktop/MSU SSQDA/SSC 442/SSC 442 Project 2/refugee_status.csv", na = c("-", "X", "D"))

non_countries <- c("Africa", "Asia", "Europe", "North America", "Oceania",
                   "South America", "Unknown", "Other", "Total")

refugees_clean <- refugees_raw %>%
  # Make this column name easier to work with
  rename(origin_country = `Continent/Country of Nationality`) %>%
  # Get rid of non-countries
  filter(!(origin_country %in% non_countries)) %>%
  # Convert country names to ISO3 codes
  mutate(iso3 = countrycode(origin_country, "country.name", "iso3c",
                            custom_match = c("Korea, North" = "PRK"))) %>%
  # Convert ISO3 codes to country names, regions, and continents
  mutate(origin_country = countrycode(iso3, "iso3c", "country.name"),
         origin_region = countrycode(iso3, "iso3c", "region"),
         origin_continent = countrycode(iso3, "iso3c", "continent")) %>%
  # Make this data tidy
  gather(year, number, -origin_country, -iso3, -origin_region, -origin_continent) %>%
  # Make sure the year column is numeric + make an actual date column for years
  mutate(year = as.numeric(year),
         year_date = ymd(paste0(year, "-01-01")))

## Cumulative country totals over time - change na's to be useful
refugees_countries_cumulative <- refugees_clean %>%
  arrange(origin_country, year) %>%
  group_by(origin_country) %>%
  mutate(cumulative_total = cumsum(number),
         cum_tot_lag1 = (cumulative_total - number),
         rate = round(number/cum_tot_lag1,4), 
         rate = ifelse(cum_tot_lag1 > 0, rate, 0)) %>%
  arrange(desc(rate), origin_country, year)

# Get total refugees from all countries to us each year
total_to_us <- refugees_clean %>%
  group_by(year) %>%
  summarize(avg_num = mean(number, na.rm = TRUE), #avg number/country in year
            ustotal = sum(number, na.rm = TRUE)) %>% #total accepted, all countries in year
  mutate(cumulative = cumsum(ustotal))

# Get percent of total refugees to us in given year for each country
p_total_us <- refugees_clean %>%
  group_by(origin_country, year) %>%
  mutate(total = sum(number, na.rm = TRUE)) %>%
  group_by(origin_country, year) %>%
  mutate(cumulative_total = cumsum(total),
         p_us_total = total/total_to_us$ustotal[total_to_us$year == year] * 100) %>%
  arrange(year, desc(p_us_total))
```

### Plots

```{r 3, echo=FALSE}
scico_col <- scico(10, palette = 'roma')

props_plot <- p_total_us %>% 
  group_by(year) %>% 
  slice_max(p_us_total, n = 5) %>%
  ggplot(aes(x = year_date, y = p_us_total, fill = origin_country)) + 
  geom_bar(position = 'stack', stat = 'identity') +
  scale_x_date(date_breaks = '1 year', date_labels = "%Y") +
  labs(x = "Year", y = "Percent of Total",
       title = "Top 5 Origin Countries for Refugees by Year",
       subtitle = "Proportion of total refugees admitted to United States by origin country, 2006 - 2015",
       caption = "Note: No annual number available for Bhutan, 2007") + # 
  theme_bw() +
  theme(plot.title = element_text(color="black", size=18, face="bold",family = "sans", hjust = 0.5),
        plot.subtitle = element_text(face = "italic", family = "serif", hjust = 0.5),
        plot.caption = element_text(hjust = 1, face = 'italic'),
        axis.title.x = element_text(color="#011111", size=14),
        axis.title.y = element_text(color="#011111", size=14),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        legend.title = element_blank(),
        legend.position = "bottom",
        axis.line = element_line(color = 'black', size = 1),
        panel.border = element_blank(),
        text = element_text(family = "sans")) +
  scale_fill_manual(values = scico_col) + 
  scale_colour_manual(values = scico_col)
props_plot

ggsave("plot.png", width = 8, height = 6)

```

## Part 2

1. What are the factors related to CO2? Whether CO2 emissions from gaseous fuel consumption and CO2 emissions from solid fuel consumption have a positive relationship with CO2 emissions?

2. How do world bank CPIA ratings correlate with a country’s ghg emissions; e.g. do countries with higher CPIA social protection ratings have greater renewable energy consumption relative to their total final energy consumption?

3. Can we use this data to model the gaps between pledges and implementation of climate goals made by major contributing countries under the Paris Agreement?

4. Does an increase in FDI into a country reduce that country’s share of non renewable to renewable energy output?

5. Do carbon dioxide emissions have a more or less  significant value in populous countries?
